---
title: "Chapter 2: Biogeography of microeukaryotes in anoxic and hypersaline polar freshwater"
format: html
editor: visual
code-fold: true
---

### Notes/unused script

```{r}
#Identify ASVs with confidence ≤ 0.97
#low_confidence_indices<-which(as.numeric(PRJNA329332_V4_taxonomy_df$Confidence)<=0.97)
#Replace taxonomic assignments with "Unidentified" for those ASVs
#taxonomy_columns<-setdiff(colnames(PRJNA329332_V4_taxonomy_df),"Confidence")
#PRJNA329332_V4_taxonomy_df[low_confidence_indices,taxonomy_columns]<-"Unidentified"
#Convert to matrix for phyloseq compatibility
#PRJNA329332_V4_taxonomy_97<-as.matrix(PRJNA329332_V4_taxonomy_df)
#Count number of unidentified sequences...
```

## Set working directory

```{r}
#Set working directory
setwd("C:/Users/ac1552/OneDrive - Natural History Museum/Documents/PhD/Biogeography/R/Biogeography")
```

## Installing and loading packages

```{r}
#Load libraries
#install.packages("ggplot2")
require(ggplot2)
#install.packages("ggforce")
require(ggforce)
#install.packages("ggtree")
require(ggtree)
#if (!require("BiocManager", quietly = TRUE))
#  install.packages("BiocManager")
#if(!requireNamespace("BiocManager")){
#  install.packages("BiocManager")
#BiocManager::install("phyloseq",force=T)
#BiocManager::install("XVector")
#BiocManager::install("UCSC.utils")
require(phyloseq)
require(tidyr)
#install.packages("forcats")
require(forcats)
require(plyr)
require(dplyr)
#for rownames_to_column() function
require(tibble)
#for write_tsv() function
require(readr)
#BiocManager::install("microbiome")
require(microbiome)
#install.packages("vegan")
require(vegan)
#install.packages("labdsv")
require(labdsv)
#install.packages("devtools")
#devtools::install_github("jiho/chroma")
require(chroma)
#install.packages("RColorBrewer")
require(RColorBrewer)
#install.packages("flextable")
require(flextable)
#install.packages("lme4")
#install.packages("lmtest")
require(lme4)
require(lmtest)
require(stringr)
require(grid)
require(scales)
require(gridExtra)
#install.packages("FSA")
require(FSA)
#install.packages("phytools")
require(phytools)
#install.packages("devtools")
#devtools::install_github("jfq3/QsRutils", build_vignettes = TRUE)
require(QsRutils)
#install.packages("ggrepel")
require(ggrepel)
#install.packages("Polychrome")
require(Polychrome)
#install.packages("remotes")
#BiocManager::install("ggtreeExtra")
require(ggtreeExtra)
#install.packages("ggnewscale")
require(ggnewscale)
```

## Set functions

```{r}
#Functions
#Prune 0 abundance records and samples from a PhyloSeq object
prune_ps<-function(ps){
  ps<-prune_samples(sample_sums(ps)>0,ps)
  ps<-prune_taxa(taxa_sums(ps)>0,ps)
  return(ps)
}

#Prune 0 abundance records from dataframe 
prune_0<-function(df){
  df<-df%>%filter(df$Abundance>0)
}

#18S Processing function
#PhyloSeq object input and glom mode
process_ps_18S<-function(ps,glom="tip",transformation="none",remove_na=T){
  #Remove 0 count taxa from PhyloSeq object
  ps<-prune_ps(ps)
  #If division agglomeration requested, otherwise do tipglom
  if(remove_na==F){
      if(glom=="Domain"){
      ps<-tax_glom(ps,taxrank="Domain",NArm=F)
    } else if(glom=="Supergroup"){
      ps<-tax_glom(ps,taxrank="Supergroup",NArm=F)
    } else if(glom=="Division"){
      ps<-tax_glom(ps,taxrank="Division",NArm=F)
    } else if(glom=="Subdivision"){
      ps<-tax_glom(ps,taxrank="Subdivision",NArm=F)
    } else if(glom=="Class"){
      ps<-tax_glom(ps,taxrank="Class",NArm=F)
    } else if(glom=="Order"){
      ps<-tax_glom(ps,taxrank="Order",NArm=F)
    } else if(glom=="Family"){
      ps<-tax_glom(ps,taxrank="Family",NArm=F)
    } else if(glom=="Genus"){
      ps<-tax_glom(ps,taxrank="Genus",NArm=F)
    } else if(glom=="Species"){
      ps<-tax_glom(ps,taxrank="Species",NArm=F)
    } else if(glom=="tip"){
      ps<-tip_glom(ps,h=0.001)
    }
  }
  else if(remove_na==T){
      if(glom=="Domain"){
      ps<-tax_glom(ps,taxrank="Domain")
    } else if(glom=="Supergroup"){
      ps<-tax_glom(ps,taxrank="Supergroup")
    } else if(glom=="Division"){
      ps<-tax_glom(ps,taxrank="Division")
    } else if(glom=="Subdivision"){
      ps<-tax_glom(ps,taxrank="Subdivision")
    } else if(glom=="Class"){
      ps<-tax_glom(ps,taxrank="Class")
    } else if(glom=="Order"){
      ps<-tax_glom(ps,taxrank="Order")
    } else if(glom=="Family"){
      ps<-tax_glom(ps,taxrank="Family")
    } else if(glom=="Genus"){
      ps<-tax_glom(ps,taxrank="Genus")
    } else if(glom=="Species"){
      ps<-tax_glom(ps,taxrank="Species")
    } else if(glom=="tip"){
      ps<-tip_glom(ps,h=0.01)
    }
  }
    #0.2 is the default value, 0.01 is the value which still retained quite a lot of info
  #If transformation is comp, then if pa, otherwise nothing
  if(transformation=="comp"){
    ps<-microbiome::transform(ps,"compositional")
  } else if(transformation=="pa"){
    ps<-microbiome::transform(ps,"pa")
  } else if(transformation=="none"){
    ps
  }
  return(ps)
}

#Theme function
theme_mine <- function(base_size=12,base_family="") {
  # Starts with theme_grey and then modify some parts
  theme_bw(base_size=base_size,base_family=base_family,
           base_line_size=base_size/24,base_rect_size=base_size/24)%+replace%
    theme(
      strip.background = element_blank(),
      strip.text.x = element_text(size = 18),
      strip.text.y = element_text(size = 18),
      axis.text.x = element_text(size=14),
      axis.text.y = element_text(size=14,hjust=1),
      axis.ticks =  element_line(colour = "black"), 
      axis.title.x= element_text(size=16),
      axis.title.y= element_text(size=16,angle=90),
      #legend.position = "none", 
      panel.background = element_blank(), 
      panel.border =element_blank(), 
      #panel.grid.major = element_blank(), 
      #panel.grid.minor = element_blank(), 
      panel.margin = unit(1.0, "lines"), 
      plot.background = element_blank(), 
      plot.margin = unit(c(0.5,  0.5, 0.5, 0.5), "lines"),
      axis.line = element_line(colour = "black")
    )
}

#Bubble theme
theme_bubble <- function(base_size=12,base_family="") {
  # Starts with theme_grey and then modify some parts
  theme_bw(base_size=base_size,base_family=base_family,
           base_line_size=base_size/24,base_rect_size=base_size/24)%+replace%
    theme(
      strip.background = element_blank(),
      strip.text.x = element_text(size = 18),
      strip.text.y = element_text(size = 18),
      axis.text.x = element_text(size=14,angle=90),
      axis.text.y = element_text(size=14,hjust=1),
      axis.ticks =  element_line(colour = "black"), 
      axis.title.x= element_text(size=16),
      axis.title.y= element_text(size=16,angle=90),
      #legend.position = "none", 
      panel.background = element_blank(), 
      panel.border =element_blank(), 
      #panel.grid.major = element_blank(), 
      #panel.grid.minor = element_blank(), 
      panel.margin = unit(1.0, "lines"), 
      plot.background = element_blank(), 
      plot.margin = unit(c(0.5,  0.5, 0.5, 0.5), "lines"),
      axis.line = element_line(colour = "black")
    )
}

remove_phyloseq_tree <- function(ps) {
  phyloseq::phyloseq(
    otu_table(ps),
    tax_table(ps),
    sample_data(ps)
  )
}
```

## Importing PRJNAs as separate phyloseq objects

### First: V4 datasets

### PRJNA329332

```{r}
#Read in OTU table
PRJNA329332_V4_otu_table=read.csv("data/PRJNA329332/V4_18S_PRJNA329332_table.csv",sep=",",row.names=1,na.strings="")
PRJNA329332_V4_otu_table=as.matrix(PRJNA329332_V4_otu_table)

#Read in taxonomy
PRJNA329332_V4_taxonomy=read.csv("data/PRJNA329332/V4_18S-PRJNA329332-pr2-taxonomy.csv",sep=",",row.names=1,na.strings="")
PRJNA329332_V4_taxonomy_df<-as.data.frame(PRJNA329332_V4_taxonomy)
#Filter out low confidence assignments
PRJNA329332_V4_taxonomy_97_df<-PRJNA329332_V4_taxonomy_df%>%filter(as.numeric(Confidence)>=0.97)
PRJNA329332_V4_taxonomy_97=as.matrix(PRJNA329332_V4_taxonomy_97_df)

#Read in metadata
#V9_18S
PRJNA329332_V4_metadata=read.csv("data/PRJNA329332/V4_18S_PRJNA329332_metadata.csv",row.names=1,na.strings="")

#Import as PhyloSeq objects
OTU_PRJNA329332_V4=otu_table(PRJNA329332_V4_otu_table,taxa_are_rows=TRUE)
TAX_PRJNA329332_V4=tax_table(PRJNA329332_V4_taxonomy_97)
META_PRJNA329332_V4=sample_data(PRJNA329332_V4_metadata)

#Check sample names are identical and same order
sample_names(OTU_PRJNA329332_V4)
sample_names(META_PRJNA329332_V4)

#Merge into one PhyloSeq object
ps_PRJNA329332_V4=phyloseq(OTU_PRJNA329332_V4,TAX_PRJNA329332_V4,META_PRJNA329332_V4)
ps_PRJNA329332_V4

#Filter out any ASVs not assigned further than 'Eukaryota' by removing any with 'NA' in the next taxonomic rank (Supergroup)
keep_taxa<-taxa_names(ps_PRJNA329332_V4)[!is.na(tax_table(ps_PRJNA329332_V4)[,"Supergroup"])]
ps_PRJNA329332_V4<-prune_taxa(keep_taxa,ps_PRJNA329332_V4)

#Add prefix to front of ASVs in ASV table and taxonomy table
prefix<-"PRJNA329332_"
old_names<-taxa_names(ps_PRJNA329332_V4)
new_names<-paste0(prefix,old_names)
taxa_names(ps_PRJNA329332_V4)<-new_names
otu_table(ps_PRJNA329332_V4)<-otu_table(ps_PRJNA329332_V4)[new_names,]
tax_table(ps_PRJNA329332_V4)<-tax_table(ps_PRJNA329332_V4)[new_names,]
```

### PRJNA803180

```{r}
#Read in OTU table
PRJNA803180_V4_otu_table=read.csv("data/PRJNA803180/V4_18S_PRJNA803180_table.csv",sep=",",row.names=1,na.strings="")
PRJNA803180_V4_otu_table=as.matrix(PRJNA803180_V4_otu_table)

#Read in taxonomy
PRJNA803180_V4_taxonomy=read.csv("data/PRJNA803180/V4_18S-PRJNA803180-pr2-taxonomy.csv",sep=",",row.names=1,na.strings="")
PRJNA803180_V4_taxonomy_df<-as.data.frame(PRJNA803180_V4_taxonomy)
#Filter out low confidence assignments
PRJNA803180_V4_taxonomy_97_df<-PRJNA803180_V4_taxonomy_df%>%filter(as.numeric(Confidence)>=0.97)
PRJNA803180_V4_taxonomy_97=as.matrix(PRJNA803180_V4_taxonomy_97_df)

#Read in metadata
#V9_18S
PRJNA803180_V4_metadata=read.csv("data/PRJNA803180/V4_18S_PRJNA803180_metadata.csv",row.names=1,na.strings="")

#Import as PhyloSeq objects
OTU_PRJNA803180_V4=otu_table(PRJNA803180_V4_otu_table,taxa_are_rows=TRUE)
TAX_PRJNA803180_V4=tax_table(PRJNA803180_V4_taxonomy_97)
META_PRJNA803180_V4=sample_data(PRJNA803180_V4_metadata)

#Check sample names are identical and same order
sample_names(OTU_PRJNA803180_V4)
sample_names(META_PRJNA803180_V4)

#Merge into one PhyloSeq object
ps_PRJNA803180_V4=phyloseq(OTU_PRJNA803180_V4,TAX_PRJNA803180_V4,META_PRJNA803180_V4)
ps_PRJNA803180_V4

#Filter out any ASVs not assigned further than 'Eukaryota' by removing any with 'NA' in the next taxonomic rank (Supergroup)
keep_taxa<-taxa_names(ps_PRJNA803180_V4)[!is.na(tax_table(ps_PRJNA803180_V4)[,"Supergroup"])]
ps_PRJNA803180_V4<-prune_taxa(keep_taxa,ps_PRJNA803180_V4)

#Add prefix to front of ASVs in ASV table and taxonomy table
prefix<-"PRJNA803180_"
old_names<-taxa_names(ps_PRJNA803180_V4)
new_names<-paste0(prefix,old_names)
taxa_names(ps_PRJNA803180_V4)<-new_names
otu_table(ps_PRJNA803180_V4)<-otu_table(ps_PRJNA803180_V4)[new_names,]
tax_table(ps_PRJNA803180_V4)<-tax_table(ps_PRJNA803180_V4)[new_names,]
```

### PRJNA805287

```{r}
#Read in OTU table
PRJNA805287_V4_otu_table=read.csv("data/PRJNA805287/V4_18S_PRJNA805287_table.csv",sep=",",row.names=1,na.strings="")
PRJNA805287_V4_otu_table=as.matrix(PRJNA805287_V4_otu_table)

#Read in taxonomy
PRJNA805287_V4_taxonomy=read.csv("data/PRJNA805287/V4_18S-PRJNA805287-pr2-taxonomy.csv",sep=",",row.names=1,na.strings="")
PRJNA805287_V4_taxonomy_df<-as.data.frame(PRJNA805287_V4_taxonomy)
#Filter out low confidence assignments
PRJNA805287_V4_taxonomy_97_df<-PRJNA805287_V4_taxonomy_df%>%filter(as.numeric(Confidence)>=0.97)
PRJNA805287_V4_taxonomy_97=as.matrix(PRJNA805287_V4_taxonomy_97_df)

#Read in metadata
#V9_18S
PRJNA805287_V4_metadata=read.csv("data/PRJNA805287/V4_18S_PRJNA805287_metadata.csv",row.names=1,na.strings="")

#Import as PhyloSeq objects
OTU_PRJNA805287_V4=otu_table(PRJNA805287_V4_otu_table,taxa_are_rows=TRUE)
TAX_PRJNA805287_V4=tax_table(PRJNA805287_V4_taxonomy_97)
META_PRJNA805287_V4=sample_data(PRJNA805287_V4_metadata)

#Check sample names are identical and same order
sample_names(OTU_PRJNA805287_V4)
sample_names(META_PRJNA805287_V4)

#Merge into one PhyloSeq object
ps_PRJNA805287_V4=phyloseq(OTU_PRJNA805287_V4,TAX_PRJNA805287_V4,META_PRJNA805287_V4)
ps_PRJNA805287_V4

#Filter out any ASVs not assigned further than 'Eukaryota' by removing any with 'NA' in the next taxonomic rank (Supergroup)
keep_taxa<-taxa_names(ps_PRJNA805287_V4)[!is.na(tax_table(ps_PRJNA805287_V4)[,"Supergroup"])]
ps_PRJNA805287_V4<-prune_taxa(keep_taxa,ps_PRJNA805287_V4)

#Add prefix to front of ASVs in ASV table and taxonomy table
prefix<-"PRJNA805287_"
old_names<-taxa_names(ps_PRJNA805287_V4)
new_names<-paste0(prefix,old_names)
taxa_names(ps_PRJNA805287_V4)<-new_names
otu_table(ps_PRJNA805287_V4)<-otu_table(ps_PRJNA805287_V4)[new_names,]
tax_table(ps_PRJNA805287_V4)<-tax_table(ps_PRJNA805287_V4)[new_names,]
```

### PRJNA838732

```{r}
#Read in OTU table
PRJNA838732_V4_otu_table=read.csv("data/PRJNA838732/V4_18S_PRJNA838732_table.csv",sep=",",row.names=1,na.strings="")
PRJNA838732_V4_otu_table=as.matrix(PRJNA838732_V4_otu_table)

#Read in taxonomy
PRJNA838732_V4_taxonomy=read.csv("data/PRJNA838732/V4_18S-PRJNA838732-pr2-taxonomy.csv",sep=",",row.names=1,na.strings="")
PRJNA838732_V4_taxonomy_df<-as.data.frame(PRJNA838732_V4_taxonomy)
#Filter out low confidence assignments
PRJNA838732_V4_taxonomy_97_df<-PRJNA838732_V4_taxonomy_df%>%filter(as.numeric(Confidence)>=0.97)
PRJNA838732_V4_taxonomy_97=as.matrix(PRJNA838732_V4_taxonomy_97_df)

#Read in metadata
#V9_18S
PRJNA838732_V4_metadata=read.csv("data/PRJNA838732/V4_18S_PRJNA838732_metadata.csv",row.names=1,na.strings="")

#Import as PhyloSeq objects
OTU_PRJNA838732_V4=otu_table(PRJNA838732_V4_otu_table,taxa_are_rows=TRUE)
TAX_PRJNA838732_V4=tax_table(PRJNA838732_V4_taxonomy_97)
META_PRJNA838732_V4=sample_data(PRJNA838732_V4_metadata)

#Check sample names are identical and same order
sample_names(OTU_PRJNA838732_V4)
sample_names(META_PRJNA838732_V4)

#Merge into one PhyloSeq object
ps_PRJNA838732_V4=phyloseq(OTU_PRJNA838732_V4,TAX_PRJNA838732_V4,META_PRJNA838732_V4)
ps_PRJNA838732_V4

#Filter out any ASVs not assigned further than 'Eukaryota' by removing any with 'NA' in the next taxonomic rank (Supergroup)
keep_taxa<-taxa_names(ps_PRJNA838732_V4)[!is.na(tax_table(ps_PRJNA838732_V4)[,"Supergroup"])]
ps_PRJNA838732_V4<-prune_taxa(keep_taxa,ps_PRJNA838732_V4)

#Add prefix to front of ASVs in ASV table and taxonomy table
prefix<-"PRJNA838732_"
old_names<-taxa_names(ps_PRJNA838732_V4)
new_names<-paste0(prefix,old_names)
taxa_names(ps_PRJNA838732_V4)<-new_names
otu_table(ps_PRJNA838732_V4)<-otu_table(ps_PRJNA838732_V4)[new_names,]
tax_table(ps_PRJNA838732_V4)<-tax_table(ps_PRJNA838732_V4)[new_names,]
```

### Next, V9 datasets

### MSc_Bonney_East

```{r}
#Read in OTU table
MSc_Bonney_East_V9_otu_table=read.csv("data/MSc_Bonney_East/V9_18S_MSc_Bonney_East_table.csv",sep=",",row.names=1,na.strings="")
MSc_Bonney_East_V9_otu_table=as.matrix(MSc_Bonney_East_V9_otu_table)

#Read in taxonomy
MSc_Bonney_East_V9_taxonomy=read.csv("data/MSc_Bonney_East/V9_18S_MSc_Bonney_East-pr2-taxonomy.csv",sep=",",row.names=1,na.strings="")
MSc_Bonney_East_V9_taxonomy_df<-as.data.frame(MSc_Bonney_East_V9_taxonomy)
#Filter out low confidence assignments
MSc_Bonney_East_V9_taxonomy_97_df<-MSc_Bonney_East_V9_taxonomy_df%>%filter(as.numeric(Confidence)>=0.97)
MSc_Bonney_East_V9_taxonomy_97=as.matrix(MSc_Bonney_East_V9_taxonomy_97_df)

#Read in metadata
#V9_18S
MSc_Bonney_East_V9_metadata=read.csv("data/MSc_Bonney_East/V9_18S_MSc_Bonney_East_metadata.csv",row.names=1,na.strings="")

#Import as PhyloSeq objects
OTU_MSc_Bonney_East_V9=otu_table(MSc_Bonney_East_V9_otu_table,taxa_are_rows=TRUE)
TAX_MSc_Bonney_East_V9=tax_table(MSc_Bonney_East_V9_taxonomy_97)
META_MSc_Bonney_East_V9=sample_data(MSc_Bonney_East_V9_metadata)

#Check sample names are identical and same order
sample_names(OTU_MSc_Bonney_East_V9)
sample_names(META_MSc_Bonney_East_V9)

#Merge into one PhyloSeq object
ps_MSc_Bonney_East_V9=phyloseq(OTU_MSc_Bonney_East_V9,TAX_MSc_Bonney_East_V9,META_MSc_Bonney_East_V9)
ps_MSc_Bonney_East_V9

#Filter out any ASVs not assigned further than 'Eukaryota' by removing any with 'NA' in the next taxonomic rank (Supergroup)
keep_taxa<-taxa_names(ps_MSc_Bonney_East_V9)[!is.na(tax_table(ps_MSc_Bonney_East_V9)[,"Supergroup"])]
ps_MSc_Bonney_East_V9<-prune_taxa(keep_taxa,ps_MSc_Bonney_East_V9)

#Add prefix to front of ASVs in ASV table and taxonomy table
prefix<-"MSc_Bonney_East_"
old_names<-taxa_names(ps_MSc_Bonney_East_V9)
new_names<-paste0(prefix,old_names)
taxa_names(ps_MSc_Bonney_East_V9)<-new_names
otu_table(ps_MSc_Bonney_East_V9)<-otu_table(ps_MSc_Bonney_East_V9)[new_names,]
tax_table(ps_MSc_Bonney_East_V9)<-tax_table(ps_MSc_Bonney_East_V9)[new_names,]
```

### MSc_Joyce

```{r}
#Read in OTU table
MSc_Joyce_V9_otu_table=read.csv("data/MSc_Joyce/V9_18S_MSc_Joyce_table.csv",sep=",",row.names=1,na.strings="")
MSc_Joyce_V9_otu_table=as.matrix(MSc_Joyce_V9_otu_table)

#Read in taxonomy
MSc_Joyce_V9_taxonomy=read.csv("data/MSc_Joyce/V9_18S_MSc_Joyce-pr2-taxonomy.csv",sep=",",row.names=1,na.strings="")
MSc_Joyce_V9_taxonomy_df<-as.data.frame(MSc_Joyce_V9_taxonomy)
#Filter out low confidence assignments
MSc_Joyce_V9_taxonomy_97_df<-MSc_Joyce_V9_taxonomy_df%>%filter(as.numeric(Confidence)>=0.97)
MSc_Joyce_V9_taxonomy_97=as.matrix(MSc_Joyce_V9_taxonomy_97_df)

#Read in metadata
#V9_18S
MSc_Joyce_V9_metadata=read.csv("data/MSc_Joyce/V9_18S_MSc_Joyce_metadata.csv",row.names=1,na.strings="")

#Import as PhyloSeq objects
OTU_MSc_Joyce_V9=otu_table(MSc_Joyce_V9_otu_table,taxa_are_rows=TRUE)
TAX_MSc_Joyce_V9=tax_table(MSc_Joyce_V9_taxonomy_97)
META_MSc_Joyce_V9=sample_data(MSc_Joyce_V9_metadata)

#Check sample names are identical and same order
sample_names(OTU_MSc_Joyce_V9)
sample_names(META_MSc_Joyce_V9)

#Merge into one PhyloSeq object
ps_MSc_Joyce_V9=phyloseq(OTU_MSc_Joyce_V9,TAX_MSc_Joyce_V9,META_MSc_Joyce_V9)
ps_MSc_Joyce_V9

#Filter out any ASVs not assigned further than 'Eukaryota' by removing any with 'NA' in the next taxonomic rank (Supergroup)
keep_taxa<-taxa_names(ps_MSc_Joyce_V9)[!is.na(tax_table(ps_MSc_Joyce_V9)[,"Supergroup"])]
ps_MSc_Joyce_V9<-prune_taxa(keep_taxa,ps_MSc_Joyce_V9)

#Add prefix to front of ASVs in ASV table and taxonomy table
prefix<-"MSc_Joyce_"
old_names<-taxa_names(ps_MSc_Joyce_V9)
new_names<-paste0(prefix,old_names)
taxa_names(ps_MSc_Joyce_V9)<-new_names
otu_table(ps_MSc_Joyce_V9)<-otu_table(ps_MSc_Joyce_V9)[new_names,]
tax_table(ps_MSc_Joyce_V9)<-tax_table(ps_MSc_Joyce_V9)[new_names,]
```

### MSc_Vanda

```{r}
#Read in OTU table
MSc_Vanda_V9_otu_table=read.csv("data/MSc_Vanda/V9_18S_MSc_Vanda_table.csv",sep=",",row.names=1,na.strings="")
MSc_Vanda_V9_otu_table=as.matrix(MSc_Vanda_V9_otu_table)

#Read in taxonomy
MSc_Vanda_V9_taxonomy=read.csv("data/MSc_Vanda/V9_18S_MSc_Vanda-pr2-taxonomy.csv",sep=",",row.names=1,na.strings="")
MSc_Vanda_V9_taxonomy_df<-as.data.frame(MSc_Vanda_V9_taxonomy)
#Filter out low confidence assignments
MSc_Vanda_V9_taxonomy_97_df<-MSc_Vanda_V9_taxonomy_df%>%filter(as.numeric(Confidence)>=0.97)
MSc_Vanda_V9_taxonomy_97=as.matrix(MSc_Vanda_V9_taxonomy_97_df)

#Read in metadata
#V9_18S
MSc_Vanda_V9_metadata=read.csv("data/MSc_Vanda/V9_18S_MSc_Vanda_metadata.csv",row.names=1,na.strings="")

#Import as PhyloSeq objects
OTU_MSc_Vanda_V9=otu_table(MSc_Vanda_V9_otu_table,taxa_are_rows=TRUE)
TAX_MSc_Vanda_V9=tax_table(MSc_Vanda_V9_taxonomy_97)
META_MSc_Vanda_V9=sample_data(MSc_Vanda_V9_metadata)

#Check sample names are identical and same order
sample_names(OTU_MSc_Vanda_V9)
sample_names(META_MSc_Vanda_V9)

#Merge into one PhyloSeq object
ps_MSc_Vanda_V9=phyloseq(OTU_MSc_Vanda_V9,TAX_MSc_Vanda_V9,META_MSc_Vanda_V9)
ps_MSc_Vanda_V9

#Filter out any ASVs not assigned further than 'Eukaryota' by removing any with 'NA' in the next taxonomic rank (Supergroup)
keep_taxa<-taxa_names(ps_MSc_Vanda_V9)[!is.na(tax_table(ps_MSc_Vanda_V9)[,"Supergroup"])]
ps_MSc_Vanda_V9<-prune_taxa(keep_taxa,ps_MSc_Vanda_V9)

#Add prefix to front of ASVs in ASV table and taxonomy table
prefix<-"MSc_Vanda_"
old_names<-taxa_names(ps_MSc_Vanda_V9)
new_names<-paste0(prefix,old_names)
taxa_names(ps_MSc_Vanda_V9)<-new_names
otu_table(ps_MSc_Vanda_V9)<-otu_table(ps_MSc_Vanda_V9)[new_names,]
tax_table(ps_MSc_Vanda_V9)<-tax_table(ps_MSc_Vanda_V9)[new_names,]
```

### PRJNA744712

```{r}
#Read in OTU table
PRJNA744712_V9_otu_table=read.csv("data/PRJNA744712/V9_18S_PRJNA744712_table.csv",sep=",",row.names=1,na.strings="")
PRJNA744712_V9_otu_table=as.matrix(PRJNA744712_V9_otu_table)

#Read in taxonomy
PRJNA744712_V9_taxonomy=read.csv("data/PRJNA744712/V9_18S_PRJNA744712-pr2-taxonomy.csv",sep=",",row.names=1,na.strings="")
PRJNA744712_V9_taxonomy_df<-as.data.frame(PRJNA744712_V9_taxonomy)
#Filter out low confidence assignments
PRJNA744712_V9_taxonomy_97_df<-PRJNA744712_V9_taxonomy_df%>%filter(as.numeric(Confidence)>=0.97)
PRJNA744712_V9_taxonomy_97=as.matrix(PRJNA744712_V9_taxonomy_97_df)

#Read in metadata
#V9_18S
PRJNA744712_V9_metadata=read.csv("data/PRJNA744712/V9_18S_PRJNA744712_metadata.csv",row.names=1,na.strings="")

#Import as PhyloSeq objects
OTU_PRJNA744712_V9=otu_table(PRJNA744712_V9_otu_table,taxa_are_rows=TRUE)
TAX_PRJNA744712_V9=tax_table(PRJNA744712_V9_taxonomy_97)
META_PRJNA744712_V9=sample_data(PRJNA744712_V9_metadata)

#Check sample names are identical and same order
sample_names(OTU_PRJNA744712_V9)
sample_names(META_PRJNA744712_V9)

#Merge into one PhyloSeq object
ps_PRJNA744712_V9=phyloseq(OTU_PRJNA744712_V9,TAX_PRJNA744712_V9,META_PRJNA744712_V9)
ps_PRJNA744712_V9

#Filter out any ASVs not assigned further than 'Eukaryota' by removing any with 'NA' in the next taxonomic rank (Supergroup)
keep_taxa<-taxa_names(ps_PRJNA744712_V9)[!is.na(tax_table(ps_PRJNA744712_V9)[,"Supergroup"])]
ps_PRJNA744712_V9<-prune_taxa(keep_taxa,ps_PRJNA744712_V9)

#Add prefix to front of ASVs in ASV table and taxonomy table
prefix<-"PRJNA744712_"
old_names<-taxa_names(ps_PRJNA744712_V9)
new_names<-paste0(prefix,old_names)
taxa_names(ps_PRJNA744712_V9)<-new_names
otu_table(ps_PRJNA744712_V9)<-otu_table(ps_PRJNA744712_V9)[new_names,]
tax_table(ps_PRJNA744712_V9)<-tax_table(ps_PRJNA744712_V9)[new_names,]
```

## Merging individual V4 phyloseq objects together

```{r}
#Merge phyloseq objects
ps_merged_V4<-merge_phyloseq(ps_PRJNA329332_V4,ps_PRJNA803180_V4,ps_PRJNA805287_V4,ps_PRJNA838732_V4)

#How many ASVs and samples
ntaxa(ps_merged_V4)
nsamples(ps_merged_V4)

#Extract metadata and export into a csv
f<-paste(getwd(),"ps_merged_V4_metadata.csv",sep="/")
merged_metadata<-meta(ps_merged_V4)
#write.csv(merged_metadata,file=f)

#Compositional transformation
ps_merged_V4_comp<-microbiome::transform(ps_merged_V4,"compositional")
#Melt into dataframe
ps_merged_V4_comp_melt<-psmelt(ps_merged_V4_comp)

#Reorder sample names
reordering_df<-ps_merged_V4_comp_melt
reordering_df$Sample.Name<-fct_relevel(reordering_df$Sample.Name,"WH_littoral_mat_RNA_1","WH_littoral_mat_RNA_2","WH_littoral_mat_RNA_3","WH_shallow_mat_RNA_1","WH_shallow_mat_RNA_2","WH_shallow_mat_RNA_3","WH_deep_mat_RNA_1","WH_deep_mat_RNA_2","WH_deep_mat_RNA_3","WH_littoral_mat_1","WH_littoral_mat_2","WH_littoral_mat_3","WH_shallow_mat_1","WH_shallow_mat_2","WH_shallow_mat_3","WH_deep_mat_1","WH_deep_mat_2","WH_deep_mat_3","WH_snow_pack_1","WH_snow_pack_2","WH_snow_pack_3","WH_water_track_1","WH_water_track_2","WH_water_track_3","WH_lake_outlet_1","WH_lake_outlet_2","WH_lake_outlet_3","WH_epilimnion_1","WH_epilimnion_2","WH_epilimnion_3","WH_mixolimnion_1","WH_mixolimnion_2","WH_mixolimnion_3","WH_hypolimnion_1","WH_hypolimnion_2","WH_hypolimnion_3","L_RNA_small","L_RNA_large","L_DNA_small","L_DNA_large","T1_RNA_small","T1_RNA_large","T1_DNA_small","T1_DNA_large","T2_RNA_small","T2_RNA_large","T2_DNA_small","T2_DNA_large","R1_RNA_small","R1_RNA_large","R1_DNA_small","R1_DNA_large","R2_RNA_small","R2_RNA_large","R2_DNA_small","R2_DNA_large","R3_RNA_small","R3_RNA_large","R3_DNA_small","R3_DNA_large","R4_RNA_small","R4_RNA_large","R4_DNA_small","R4_DNA_large","R5_RNA_small","R5_RNA_large","R5_DNA_small","R5_DNA_large","R6_RNA_small","R6_RNA_large","R6_DNA_small","R6_DNA_large","R7_RNA_small","R7_RNA_large","R7_DNA_small","R7_DNA_large","R8_RNA_small","R8_RNA_large","R8_DNA_small","R8_DNA_large","R9_RNA_small","R9_RNA_large","R9_DNA_small","R9_DNA_large","P1_RNA_small","P1_RNA_large","P1_DNA_small","P1_DNA_large","P2_RNA_small","P2_RNA_large","P2_DNA_small","P2_DNA_large","P3_RNA_small","P3_RNA_large","P3_DNA_small","P3_DNA_large","B_RNA_small","B_RNA_large","B_DNA_small","B_DNA_large","WS_surface","KS_surface","KS_upper_0.5a","KS_upper_0.5b","KS_upper_1a","KS_upper_1b","KS_mid_1.5a","KS_mid_1.5b","KS_mid_2a","KS_mid_2b","KS_chemocline_2.5a","KS_chemocline_2.5b","KS_chemocline_3a","KS_chemocline_3b","KS_chemocline_3.1a","KS_chemocline_3.1b","KS_anoxic_3.5","KS_anoxic_4a","KS_anoxic_4b","Chang_Hu_17","Chang_Hu_18","Chang_Hu_19","Kitec_17","Kitec_18","Kitec_19","Xi_Hu_17","Xi_Hu_18","Xi_Hu_19","Yan_Ou_17","Yan_Ou_18","Yan_Ou_19","Yue_Ya_17","Yue_Ya_18","Yue_Ya_19")
reordering_df$env_broad_scale<-fct_relevel(reordering_df$env_broad_scale,"Ward_Hunt_Lake","Sheldrake_Lake","Sheldrake_River_Tributary","Sheldrake_River","Hudson_Bay_Estuary","Hudson_Bay","White Sea","Kislo-Sladkoe Lake","Chang_Hu_Lake","Kitec_Lake","Xi_Hu_Lake","Yan_Ou_Lake","Yue_Ya_Lake")
ps_merged_V4_comp_melt<-reordering_df

#Name colour palettes using this dataframe
site_palette<-unname(createPalette(length(unique(reordering_df$env_broad_scale)),c("#010101","#ffffff")))
```

## Alpha diversity

```{r}
#Calculate alpha diversity indices
merged_V4_alpha<-estimate_richness(ps_merged_V4,measures=c("Observed","Shannon","Simpson"))
#Add sample metadata to this new alpha diversity dataframe
merged_V4_alpha$SampleID<-rownames(merged_V4_alpha)
merged_V4_metadata<-as(sample_data(ps_merged_V4),"data.frame")
merged_V4_alpha<-cbind(merged_V4_alpha,merged_V4_metadata[merged_V4_alpha$SampleID,])
#reshape to long format
merged_V4_alpha_long<-pivot_longer(merged_V4_alpha,cols=c("Observed","Shannon","Simpson"),names_to="Index",values_to="Diversity")



#Reorder sample ids
reordering_df<-merged_V4_alpha_long
reordering_df$Sample.Name<-fct_relevel(reordering_df$Sample.Name,"WH_littoral_mat_RNA_1","WH_littoral_mat_RNA_2","WH_littoral_mat_RNA_3","WH_shallow_mat_RNA_1","WH_shallow_mat_RNA_2","WH_shallow_mat_RNA_3","WH_deep_mat_RNA_1","WH_deep_mat_RNA_2","WH_deep_mat_RNA_3","WH_littoral_mat_1","WH_littoral_mat_2","WH_littoral_mat_3","WH_shallow_mat_1","WH_shallow_mat_2","WH_shallow_mat_3","WH_deep_mat_1","WH_deep_mat_2","WH_deep_mat_3","WH_snow_pack_1","WH_snow_pack_2","WH_snow_pack_3","WH_water_track_1","WH_water_track_2","WH_water_track_3","WH_lake_outlet_1","WH_lake_outlet_2","WH_lake_outlet_3","WH_epilimnion_1","WH_epilimnion_2","WH_epilimnion_3","WH_mixolimnion_1","WH_mixolimnion_2","WH_mixolimnion_3","WH_hypolimnion_1","WH_hypolimnion_2","WH_hypolimnion_3","L_RNA_small","L_RNA_large","L_DNA_small","L_DNA_large","T1_RNA_small","T1_RNA_large","T1_DNA_small","T1_DNA_large","T2_RNA_small","T2_RNA_large","T2_DNA_small","T2_DNA_large","R1_RNA_small","R1_RNA_large","R1_DNA_small","R1_DNA_large","R2_RNA_small","R2_RNA_large","R2_DNA_small","R2_DNA_large","R3_RNA_small","R3_RNA_large","R3_DNA_small","R3_DNA_large","R4_RNA_small","R4_RNA_large","R4_DNA_small","R4_DNA_large","R5_RNA_small","R5_RNA_large","R5_DNA_small","R5_DNA_large","R6_RNA_small","R6_RNA_large","R6_DNA_small","R6_DNA_large","R7_RNA_small","R7_RNA_large","R7_DNA_small","R7_DNA_large","R8_RNA_small","R8_RNA_large","R8_DNA_small","R8_DNA_large","R9_RNA_small","R9_RNA_large","R9_DNA_small","R9_DNA_large","P1_RNA_small","P1_RNA_large","P1_DNA_small","P1_DNA_large","P2_RNA_small","P2_RNA_large","P2_DNA_small","P2_DNA_large","P3_RNA_small","P3_RNA_large","P3_DNA_small","P3_DNA_large","B_RNA_small","B_RNA_large","B_DNA_small","B_DNA_large","WS_surface","KS_surface","KS_upper_0.5a","KS_upper_0.5b","KS_upper_1a","KS_upper_1b","KS_mid_1.5a","KS_mid_1.5b","KS_mid_2a","KS_mid_2b","KS_chemocline_2.5a","KS_chemocline_2.5b","KS_chemocline_3a","KS_chemocline_3b","KS_chemocline_3.1a","KS_chemocline_3.1b","KS_anoxic_3.5","KS_anoxic_4a","KS_anoxic_4b","Chang_Hu_17","Chang_Hu_18","Chang_Hu_19","Kitec_17","Kitec_18","Kitec_19","Xi_Hu_17","Xi_Hu_18","Xi_Hu_19","Yan_Ou_17","Yan_Ou_18","Yan_Ou_19","Yue_Ya_17","Yue_Ya_18","Yue_Ya_19")
reordering_df$env_broad_scale<-fct_relevel(reordering_df$env_broad_scale,"Ward_Hunt_Lake","Sheldrake_Lake","Sheldrake_River_Tributary","Sheldrake_River","Hudson_Bay_Estuary","Hudson_Bay","White Sea","Kislo-Sladkoe Lake","Chang_Hu_Lake","Kitec_Lake","Xi_Hu_Lake","Yan_Ou_Lake","Yue_Ya_Lake")
merged_V4_alpha_long<-reordering_df

#colour palette
site_palette<-unname(createPalette(length(unique(reordering_df$env_broad_scale)),c("#010101","#ffffff")))

#plot
p<-ggplot(merged_V4_alpha_long,aes(x=env_broad_scale,y=Diversity,colour=env_broad_scale,fill=env_broad_scale))+
  scale_colour_manual("Broad Location",values=site_palette)+
  scale_fill_manual("Broad Location",values=site_palette)+
  geom_point(size=3)+
  geom_boxplot(alpha=0.8)+
  facet_wrap(~Index,scales="free_y")+
  labs(x="Site",y="Alpha Diversity")+
  theme(axis.text.x=element_text(angle=90,vjust=1,hjust=1))
print(p)

ggsave("plots/ps_merged_V4_alpha.svg",width=400,height=400,units=c("mm"))
```

## Beta diversity - NMDS

```{r}
#NMDS
ps_merged_V4_Class_comp<-process_ps_18S(ps_merged_V4,glom="Class",transformation="comp",remove_na=T)
ps_merged_V4_Class_comp_melt<-psmelt(ps_merged_V4_Class_comp)
ps_merged_V4_Class_comp_melt<-prune_0(ps_merged_V4_Class_comp_melt)
#Reordering
reordering_df<-ps_merged_V4_Class_comp_melt
reordering_df$Sample.Name<-fct_relevel(reordering_df$Sample.Name,"WH_littoral_mat_RNA_1","WH_littoral_mat_RNA_2","WH_littoral_mat_RNA_3","WH_shallow_mat_RNA_1","WH_shallow_mat_RNA_2","WH_shallow_mat_RNA_3","WH_deep_mat_RNA_1","WH_deep_mat_RNA_2","WH_deep_mat_RNA_3","WH_littoral_mat_1","WH_littoral_mat_2","WH_littoral_mat_3","WH_shallow_mat_1","WH_shallow_mat_2","WH_shallow_mat_3","WH_deep_mat_1","WH_deep_mat_2","WH_deep_mat_3","WH_snow_pack_1","WH_snow_pack_2","WH_snow_pack_3","WH_water_track_1","WH_water_track_2","WH_water_track_3","WH_lake_outlet_1","WH_lake_outlet_2","WH_lake_outlet_3","WH_epilimnion_1","WH_epilimnion_2","WH_epilimnion_3","WH_mixolimnion_1","WH_mixolimnion_2","WH_mixolimnion_3","WH_hypolimnion_1","WH_hypolimnion_2","WH_hypolimnion_3","L_RNA_small","L_RNA_large","L_DNA_small","L_DNA_large","T1_RNA_small","T1_RNA_large","T1_DNA_small","T1_DNA_large","T2_RNA_small","T2_RNA_large","T2_DNA_small","T2_DNA_large","R1_RNA_small","R1_RNA_large","R1_DNA_small","R1_DNA_large","R2_RNA_small","R2_RNA_large","R2_DNA_small","R2_DNA_large","R3_RNA_small","R3_RNA_large","R3_DNA_small","R3_DNA_large","R4_RNA_small","R4_RNA_large","R4_DNA_small","R4_DNA_large","R5_RNA_small","R5_RNA_large","R5_DNA_small","R5_DNA_large","R6_RNA_small","R6_RNA_large","R6_DNA_small","R6_DNA_large","R7_RNA_small","R7_RNA_large","R7_DNA_small","R7_DNA_large","R8_RNA_small","R8_RNA_large","R8_DNA_small","R8_DNA_large","R9_RNA_small","R9_RNA_large","R9_DNA_small","R9_DNA_large","P1_RNA_small","P1_RNA_large","P1_DNA_small","P1_DNA_large","P2_RNA_small","P2_RNA_large","P2_DNA_small","P2_DNA_large","P3_RNA_small","P3_RNA_large","P3_DNA_small","P3_DNA_large","B_RNA_small","B_RNA_large","B_DNA_small","B_DNA_large","WS_surface","KS_surface","KS_upper_0.5a","KS_upper_0.5b","KS_upper_1a","KS_upper_1b","KS_mid_1.5a","KS_mid_1.5b","KS_mid_2a","KS_mid_2b","KS_chemocline_2.5a","KS_chemocline_2.5b","KS_chemocline_3a","KS_chemocline_3b","KS_chemocline_3.1a","KS_chemocline_3.1b","KS_anoxic_3.5","KS_anoxic_4a","KS_anoxic_4b","Chang_Hu_17","Chang_Hu_18","Chang_Hu_19","Kitec_17","Kitec_18","Kitec_19","Xi_Hu_17","Xi_Hu_18","Xi_Hu_19","Yan_Ou_17","Yan_Ou_18","Yan_Ou_19","Yue_Ya_17","Yue_Ya_18","Yue_Ya_19")
reordering_df$env_broad_scale<-fct_relevel(reordering_df$env_broad_scale,"Ward_Hunt_Lake","Sheldrake_Lake","Sheldrake_River_Tributary","Sheldrake_River","Hudson_Bay_Estuary","Hudson_Bay","White Sea","Kislo-Sladkoe Lake","Chang_Hu_Lake","Kitec_Lake","Xi_Hu_Lake","Yan_Ou_Lake","Yue_Ya_Lake")
ps_merged_V4_Class_comp_melt<-reordering_df

#colour palette
site_palette<-unname(createPalette(length(unique(reordering_df$env_broad_scale)),c("#010101","#ffffff")))

#Ordination
ps_merged_V4_Class_comp_ord_weighted<-ordinate(ps_merged_V4_Class_comp,"NMDS","bray")
#Stress: 0.1589103
#ANOSIMs
ps_merged_V4_Class_comp_loc_group=get_variable(ps_merged_V4_Class_comp, "env_broad_scale")
ps_merged_V4_Class_comp_loc_anosim=anosim(distance(ps_merged_V4_Class_comp, "bray"), ps_merged_V4_Class_comp_loc_group)
ps_merged_V4_Class_comp_loc_anosim
#ANOSIM: Stress = 0.1819698, ANOSIM R-Stat 0.4945, p value = 0.001


ps_merged_V4_Class_comp_dataframe<-as.data.frame(sample_data(ps_merged_V4_Class_comp))
ps_merged_V4_Class_comp_nmds_weighted<-plot_ordination(ps_merged_V4_Class_comp,ps_merged_V4_Class_comp_ord_weighted,type="samples",
                                      color="env_broad_scale",label="Sample.Name",shape="env_medium")+
  ggtitle("ANOSIM: Stress = 0.1819698, ANOSIM R-Stat 0.4945, p value = 0.001")+
  scale_colour_manual("Location",values=site_palette)+
  scale_shape_manual("Media",values=c(15,16,17))+
  geom_point(size=4,alpha=0.8)+
  geom_text_repel(mapping=aes(x=NMDS1,y=NMDS2,label=Sample.Name,fontface="bold"),inherit.aes=F,,size = 3,vjust=0,nudge_y=0.04,max.overlaps=100)+
  geom_mark_hull(aes(fill=env_broad_scale,colour=env_broad_scale),show.legend=F)+
  expand_limits(x=1.5,y=1)+
  theme_mine()
ps_merged_V4_Class_comp_nmds_weighted$layers
ps_merged_V4_Class_comp_nmds_weighted$layers<-ps_merged_V4_Class_comp_nmds_weighted$layers[-1]
ps_merged_V4_Class_comp_nmds_weighted$layers<-ps_merged_V4_Class_comp_nmds_weighted$layers[-1]
ps_merged_V4_Class_comp_nmds_weighted
ggsave("Plots/ps_merged_V4_Class_comp_nmds_weighted.svg",width=400,height=250,units=c("mm"))
```

## Relative Abundance bubble plots

### Class level

```{r}
#Class level
ps_merged_V4_Class_comp<-process_ps_18S(ps_merged_V4,glom="Class",transformation="comp",remove_na=F)
#Extract taxonomy table
tax<-tax_table(ps_merged_V4_Class_comp)
unresolved_taxa<-is.na(tax[,"Class"])
#Count number of unresolved asvs
num_unresolved<-sum(unresolved_taxa)
cat("Number of ASVs not resolved to class level:",num_unresolved,"\n")
#Make a copy to edit
tax_fixed<-tax
#Make a new pseudo-class called 'Unresolved', and apply it to higher taxonomic levels
tax_fixed[unresolved_taxa,"Class"]<-"Unresolved"
tax_fixed[unresolved_taxa,"Subdivision"]<-"Unresolved"
tax_fixed[unresolved_taxa,"Division"]<-"Unresolved"
tax_fixed[unresolved_taxa,"Supergroup"]<-"Unresolved"
tax_fixed[unresolved_taxa,"Domain"]<-"Unresolved"
#Replace tax table of phyloseq object
tax_table(ps_merged_V4_Class_comp)<-tax_fixed
#Reprocess the ps object
ps_merged_V4_Class_comp<-tax_glom(ps_merged_V4_Class_comp,taxrank="Class")
ps_merged_V4_Class_comp_melt<-psmelt(ps_merged_V4_Class_comp)
ps_merged_V4_Class_comp_melt[["Class"]]<-as.character(ps_merged_V4_Class_comp_melt[["Class"]])
ps_merged_V4_Class_comp_melt<-prune_0(ps_merged_V4_Class_comp_melt)

#Reordering
reordering_df<-ps_merged_V4_Class_comp_melt
reordering_df$Sample.Name<-fct_relevel(reordering_df$Sample.Name,"WH_littoral_mat_RNA_1","WH_littoral_mat_RNA_2","WH_littoral_mat_RNA_3","WH_shallow_mat_RNA_1","WH_shallow_mat_RNA_2","WH_shallow_mat_RNA_3","WH_deep_mat_RNA_1","WH_deep_mat_RNA_2","WH_deep_mat_RNA_3","WH_littoral_mat_1","WH_littoral_mat_2","WH_littoral_mat_3","WH_shallow_mat_1","WH_shallow_mat_2","WH_shallow_mat_3","WH_deep_mat_1","WH_deep_mat_2","WH_deep_mat_3","WH_snow_pack_1","WH_snow_pack_2","WH_snow_pack_3","WH_water_track_1","WH_water_track_2","WH_water_track_3","WH_lake_outlet_1","WH_lake_outlet_2","WH_lake_outlet_3","WH_epilimnion_1","WH_epilimnion_2","WH_epilimnion_3","WH_mixolimnion_1","WH_mixolimnion_2","WH_mixolimnion_3","WH_hypolimnion_1","WH_hypolimnion_2","WH_hypolimnion_3","L_RNA_small","L_RNA_large","L_DNA_small","L_DNA_large","T1_RNA_small","T1_RNA_large","T1_DNA_small","T1_DNA_large","T2_RNA_small","T2_RNA_large","T2_DNA_small","T2_DNA_large","R1_RNA_small","R1_RNA_large","R1_DNA_small","R1_DNA_large","R2_RNA_small","R2_RNA_large","R2_DNA_small","R2_DNA_large","R3_RNA_small","R3_RNA_large","R3_DNA_small","R3_DNA_large","R4_RNA_small","R4_RNA_large","R4_DNA_small","R4_DNA_large","R5_RNA_small","R5_RNA_large","R5_DNA_small","R5_DNA_large","R6_RNA_small","R6_RNA_large","R6_DNA_small","R6_DNA_large","R7_RNA_small","R7_RNA_large","R7_DNA_small","R7_DNA_large","R8_RNA_small","R8_RNA_large","R8_DNA_small","R8_DNA_large","R9_RNA_small","R9_RNA_large","R9_DNA_small","R9_DNA_large","P1_RNA_small","P1_RNA_large","P1_DNA_small","P1_DNA_large","P2_RNA_small","P2_RNA_large","P2_DNA_small","P2_DNA_large","P3_RNA_small","P3_RNA_large","P3_DNA_small","P3_DNA_large","B_RNA_small","B_RNA_large","B_DNA_small","B_DNA_large","WS_surface","KS_surface","KS_upper_0.5a","KS_upper_0.5b","KS_upper_1a","KS_upper_1b","KS_mid_1.5a","KS_mid_1.5b","KS_mid_2a","KS_mid_2b","KS_chemocline_2.5a","KS_chemocline_2.5b","KS_chemocline_3a","KS_chemocline_3b","KS_chemocline_3.1a","KS_chemocline_3.1b","KS_anoxic_3.5","KS_anoxic_4a","KS_anoxic_4b","Chang_Hu_17","Chang_Hu_18","Chang_Hu_19","Kitec_17","Kitec_18","Kitec_19","Xi_Hu_17","Xi_Hu_18","Xi_Hu_19","Yan_Ou_17","Yan_Ou_18","Yan_Ou_19","Yue_Ya_17","Yue_Ya_18","Yue_Ya_19")
reordering_df$env_broad_scale<-fct_relevel(reordering_df$env_broad_scale,"Ward_Hunt_Lake","Sheldrake_Lake","Sheldrake_River_Tributary","Sheldrake_River","Hudson_Bay_Estuary","Hudson_Bay","White Sea","Kislo-Sladkoe Lake","Chang_Hu_Lake","Kitec_Lake","Xi_Hu_Lake","Yan_Ou_Lake","Yue_Ya_Lake")
ps_merged_V4_Class_comp_melt<-reordering_df


# Calculate mean abundance by Class
mean_abund <- ps_merged_V4_Class_comp_melt %>%
  group_by(Class) %>%
  summarise(mean_abund = mean(Abundance, na.rm = TRUE)) %>%
  arrange(desc(mean_abund))
# Get levels in abundance order
Class_levels<-mean_abund$Class
# Move 'Unresolved' to the end
Class_levels<-c(setdiff(Class_levels, "Unresolved"), "Unresolved")
# Apply new levels
ps_merged_V4_Class_comp_melt$Class<-factor(ps_merged_V4_Class_comp_melt$Class,levels=Class_levels)

#Group low abundance data together (<0.01%)
ps_merged_V4_Class_comp_melt[["Class"]][ps_merged_V4_Class_comp_melt$Abundance<0.001]<-"<0.1% total abundance"
#Remove any NAs generated
ps_merged_V4_Class_comp_melt$Class[is.na(ps_merged_V4_Class_comp_melt$Class)] <- "Unresolved"
#Palette with lots of colours
site_palette<-unname(createPalette(length(unique(reordering_df$env_broad_scale)),c("#010101","#ffffff")))

#plot bubble plot
bubble_plot<-ggplot(ps_merged_V4_Class_comp_melt,aes(shape=geo_loc_name,x=Sample.Name,y=fct_rev(Class)))+
  geom_point(aes(size=Abundance,color=as.factor(env_broad_scale),shape=env_medium),alpha=0.8)+
  scale_size_continuous("Relative Abundance",limits=c(0.00001,1),range=c(1.5,15),breaks=c(0.10,0.25,0.50,0.75,1),labels=c("10%","25%","50%","75%","100%"))+
  scale_color_manual("Location",values=site_palette)+
  scale_shape_manual("Country",values=c(16,15,17,18))+
  labs(y="Class",x="Sample")+
  theme_bubble()+
  expand_limits(y=c(0,length(levels(ps_merged_V4_Class_comp_melt$"Class"))),x=c(0,length(levels(ps_merged_V4_Class_comp_melt$"Sample.Name"))))+
  guides(size=guide_legend(order=1),colour=guide_legend(order=2))
bubble_plot$layers
bubble_plot
ggsave("Plots/ps_merged_V4_Class_bubble_plot_with_unresolved.svg",width=600,height=400,units=c("mm"))
```

## Averaging across replicates

```{r}
#Convert NAs to 0 abundance
otu_table(ps_merged_V4)[is.na(otu_table(ps_merged_V4))]<-0

#Preserve metadata
metadata_backup <- as(sample_data(ps_merged_V4_comp), "data.frame")
metadata_backup$SampleID <- rownames(metadata_backup)

#Merge samples according to their replicate group
ps_merged_V4_grouped<-merge_samples(ps_merged_V4,group="RepGroup")


#Create new metadata manually, using first row per group since the replicates had identical metadata
metadata_grouped <- metadata_backup %>%
  group_by(RepGroup) %>%
  slice(1) %>%
  ungroup()

#Force setting of sample names to correct rownames (RepGroup)
metadata_grouped_sd<-sample_data(as.data.frame(metadata_grouped))
sample_names(metadata_grouped_sd)<-metadata_grouped$RepGroup
sample_data(ps_merged_V4_grouped) <- metadata_grouped_sd


#get number of replicates per group
rep_counts<-table(sample_data(ps_merged_V4)$RepGroup)

#extract and convert otu table to a matrix
otu_mat<-as(otu_table(ps_merged_V4_grouped),"matrix")
otu_mat<-t(otu_mat)
#ensure sample names match
missing_names <- setdiff(colnames(otu_mat), names(rep_counts))
if (length(missing_names) > 0) {
  print("Warning: These merged sample names are missing from rep_counts:")
  print(missing_names)
}
colnames(otu_mat) <- as.character(colnames(otu_mat))
names(rep_counts) <- as.character(names(rep_counts))


#Average
otu_mat<-sweep(otu_mat,2,rep_counts[colnames(otu_mat)],"/")
otu_mat[is.na(otu_mat)] <- 0
#average merged abundances
#otu_table(ps_merged_V4_comp_grouped)<-otu_table(ps_merged_V4_comp_grouped)/rep_counts[rownames(otu_table(ps_merged_V4_comp_grouped))]

#Check and align sample names
otu_mat <- otu_mat[, rownames(sample_data(ps_merged_V4_grouped))]

#Return back to phyloseq object
otu_table(ps_merged_V4_grouped) <- otu_table(otu_mat, taxa_are_rows = TRUE)
```

## Relative Abundance bubble plots

### Class level

```{r}
#Class level
ps_merged_V4_grouped_Class_comp<-process_ps_18S(ps_merged_V4_grouped,glom="Class",transformation="comp",remove_na=F)
#Extract taxonomy table
tax<-tax_table(ps_merged_V4_grouped_Class_comp)
unresolved_taxa<-is.na(tax[,"Class"])
#Count number of unresolved asvs
num_unresolved<-sum(unresolved_taxa)
cat("Number of ASVs not resolved to class level:",num_unresolved,"\n")
#Make a copy to edit
tax_fixed<-tax
#Make a new pseudo-class called 'Unresolved', and apply it to higher taxonomic levels
tax_fixed[unresolved_taxa,"Class"]<-"Unresolved"
tax_fixed[unresolved_taxa,"Subdivision"]<-"Unresolved"
tax_fixed[unresolved_taxa,"Division"]<-"Unresolved"
tax_fixed[unresolved_taxa,"Supergroup"]<-"Unresolved"
tax_fixed[unresolved_taxa,"Domain"]<-"Unresolved"
#Replace tax table of phyloseq object
tax_table(ps_merged_V4_grouped_Class_comp)<-tax_fixed
#Reprocess the ps object
ps_merged_V4_grouped_Class_comp<-tax_glom(ps_merged_V4_grouped_Class_comp,taxrank="Class")
ps_merged_V4_grouped_Class_comp_melt<-psmelt(ps_merged_V4_grouped_Class_comp)
ps_merged_V4_grouped_Class_comp_melt[["Class"]]<-as.character(ps_merged_V4_grouped_Class_comp_melt[["Class"]])
ps_merged_V4_grouped_Class_comp_melt<-prune_0(ps_merged_V4_grouped_Class_comp_melt)

#Reordering
reordering_df<-ps_merged_V4_grouped_Class_comp_melt
reordering_df$RepGroup<-fct_relevel(reordering_df$RepGroup,"WH_littoral_mat","WH_shallow_mat","WH_deep_mat","WH_snow_pack","WH_water_track","WH_lake_outlet","WH_epilimnion","WH_mixolimnion","WH_hypolimnion","Sheldrake_Lake","Sheldrake_River_Tributary_1","Sheldrake_River_Tributary_2","Sheldrake_River_1","Sheldrake_River_2","Sheldrake_River_3","Sheldrake_River_4","Sheldrake_River_5","Sheldrake_River_6","Sheldrake_River_7","Sheldrake_River_8","Sheldrake_River_9","Hudson_Bay_Estuary_1","Hudson_Bay_Estuary_2","Hudson_Bay_Estuary_3","Hudson_Bay","WS_surface","KS_surface","KS_upper_0.5","KS_upper_1","KS_mid_1.5","KS_mid_2","KS_chemocline_2.5","KS_chemocline_3","KS_chemocline_3.1","KS_anoxic_3.5","KS_anoxic_4","Chang_Hu_17","Chang_Hu_18","Chang_Hu_19","Kitec_17","Kitec_18","Kitec_19","Xi_Hu_17","Xi_Hu_18","Xi_Hu_19","Yan_Ou_17","Yan_Ou_18","Yan_Ou_19","Yue_Ya_17","Yue_Ya_18","Yue_Ya_19")
reordering_df$env_broad_scale<-fct_relevel(reordering_df$env_broad_scale,"Ward_Hunt_Lake","Sheldrake_Lake","Sheldrake_River_Tributary","Sheldrake_River","Hudson_Bay_Estuary","Hudson_Bay","White Sea","Kislo-Sladkoe Lake","Chang_Hu_Lake","Kitec_Lake","Xi_Hu_Lake","Yan_Ou_Lake","Yue_Ya_Lake")
ps_merged_V4_grouped_Class_comp_melt<-reordering_df


# Calculate mean abundance by Class
mean_abund <- ps_merged_V4_grouped_Class_comp_melt %>%
  group_by(Class) %>%
  summarise(mean_abund = mean(Abundance, na.rm = TRUE)) %>%
  arrange(desc(mean_abund))
# Get levels in abundance order
Class_levels<-mean_abund$Class
# Move 'Unresolved' to the end
Class_levels<-c(setdiff(Class_levels, "Unresolved"), "Unresolved")
# Apply new levels
ps_merged_V4_grouped_Class_comp_melt$Class<-factor(ps_merged_V4_grouped_Class_comp_melt$Class,levels=Class_levels)

#Group low abundance data together (<0.01%)
ps_merged_V4_grouped_Class_comp_melt[["Class"]][ps_merged_V4_grouped_Class_comp_melt$Abundance<0.001]<-"<0.1% total abundance"
#Remove any NAs generated
ps_merged_V4_grouped_Class_comp_melt$Class[is.na(ps_merged_V4_grouped_Class_comp_melt$Class)] <- "Unresolved"
#Palette with lots of colours
site_palette<-unname(createPalette(length(unique(reordering_df$env_broad_scale)),c("#010101","#ffffff")))

#plot bubble plot
bubble_plot<-ggplot(ps_merged_V4_grouped_Class_comp_melt,aes(shape=geo_loc_name,x=RepGroup,y=fct_rev(Class)))+
  geom_point(aes(size=Abundance,color=as.factor(env_broad_scale),shape=env_medium),alpha=0.8)+
  scale_size_continuous("Relative Abundance",limits=c(0.00001,1),range=c(1.5,15),breaks=c(0.10,0.25,0.50,0.75,1),labels=c("10%","25%","50%","75%","100%"))+
  scale_color_manual("Location",values=site_palette)+
  scale_shape_manual("Country",values=c(16,15,17,18))+
  labs(y="Class",x="Sample")+
  theme_bubble()+
  #expand_limits(y=c(0,length(levels(ps_merged_V4_grouped_Class_comp_melt$"Class"))),x=c(0,length(levels(ps_merged_V4_grouped_Class_comp_melt$"RepGroup"))))+
  guides(size=guide_legend(order=1),colour=guide_legend(order=2))
bubble_plot$layers
bubble_plot
ggsave("Plots/ps_merged_V4_grouped_Class_bubble_plot_with_unresolved.svg",width=600,height=400,units=c("mm"))
```

## Beta diversity - NMDS

### Class level

```{r}
#NMDS
ps_merged_V4_grouped_Class_comp<-process_ps_18S(ps_merged_V4_grouped,glom="Class",transformation="comp",remove_na=T)
ps_merged_V4_grouped_Class_comp_melt<-psmelt(ps_merged_V4_grouped_Class_comp)
ps_merged_V4_grouped_Class_comp_melt<-prune_0(ps_merged_V4_grouped_Class_comp_melt)
#Reordering
#Reordering
reordering_df<-ps_merged_V4_grouped_Class_comp_melt
reordering_df$RepGroup<-fct_relevel(reordering_df$RepGroup,"WH_littoral_mat","WH_shallow_mat","WH_deep_mat","WH_snow_pack","WH_water_track","WH_lake_outlet","WH_epilimnion","WH_mixolimnion","WH_hypolimnion","Sheldrake_Lake","Sheldrake_River_Tributary_1","Sheldrake_River_Tributary_2","Sheldrake_River_1","Sheldrake_River_2","Sheldrake_River_3","Sheldrake_River_4","Sheldrake_River_5","Sheldrake_River_6","Sheldrake_River_7","Sheldrake_River_8","Sheldrake_River_9","Hudson_Bay_Estuary_1","Hudson_Bay_Estuary_2","Hudson_Bay_Estuary_3","Hudson_Bay","WS_surface","KS_surface","KS_upper_0.5","KS_upper_1","KS_mid_1.5","KS_mid_2","KS_chemocline_2.5","KS_chemocline_3","KS_chemocline_3.1","KS_anoxic_3.5","KS_anoxic_4","Chang_Hu_17","Chang_Hu_18","Chang_Hu_19","Kitec_17","Kitec_18","Kitec_19","Xi_Hu_17","Xi_Hu_18","Xi_Hu_19","Yan_Ou_17","Yan_Ou_18","Yan_Ou_19","Yue_Ya_17","Yue_Ya_18","Yue_Ya_19")
reordering_df$env_broad_scale<-fct_relevel(reordering_df$env_broad_scale,"Ward_Hunt_Lake","Sheldrake_Lake","Sheldrake_River_Tributary","Sheldrake_River","Hudson_Bay_Estuary","Hudson_Bay","White Sea","Kislo-Sladkoe Lake","Chang_Hu_Lake","Kitec_Lake","Xi_Hu_Lake","Yan_Ou_Lake","Yue_Ya_Lake")
ps_merged_V4_grouped_Class_comp_melt<-reordering_df

#colour palette
site_palette<-unname(createPalette(length(unique(reordering_df$env_broad_scale)),c("#010101","#ffffff")))

#Ordination
ps_merged_V4_grouped_Class_comp_ord_weighted<-ordinate(ps_merged_V4_grouped_Class_comp,"NMDS","bray")
#Stress: 0.1589103
#ANOSIMs
ps_merged_V4_grouped_Class_comp_loc_group=get_variable(ps_merged_V4_grouped_Class_comp, "env_broad_scale")
ps_merged_V4_grouped_Class_comp_loc_anosim=anosim(distance(ps_merged_V4_grouped_Class_comp, "bray"), ps_merged_V4_grouped_Class_comp_loc_group)
ps_merged_V4_grouped_Class_comp_loc_anosim
#ANOSIM: Stress = 0.1447421, ANOSIM R-Stat 0.4238, p value = 0.001


ps_merged_V4_grouped_Class_comp_dataframe<-as.data.frame(sample_data(ps_merged_V4_grouped_Class_comp))
ps_merged_V4_grouped_Class_comp_nmds_weighted<-plot_ordination(ps_merged_V4_grouped_Class_comp,ps_merged_V4_grouped_Class_comp_ord_weighted,type="samples",
                                      color="env_broad_scale",label="RepGroup",shape="env_medium")+
  ggtitle("ANOSIM: Stress = 0.1447421, ANOSIM R-Stat 0.4238, p value = 0.001")+
  scale_colour_manual("Location",values=site_palette)+
  scale_shape_manual("Media",values=c(15,16,17))+
  geom_point(size=4,alpha=0.8)+
  geom_text_repel(mapping=aes(x=NMDS1,y=NMDS2,label=RepGroup,fontface="bold"),inherit.aes=F,,size = 3,vjust=0,nudge_y=0.04,max.overlaps=100)+
  geom_mark_hull(aes(fill=env_broad_scale,colour=env_broad_scale),show.legend=F)+
  expand_limits(x=1.5,y=1)+
  theme_mine()
ps_merged_V4_grouped_Class_comp_nmds_weighted$layers
ps_merged_V4_grouped_Class_comp_nmds_weighted$layers<-ps_merged_V4_grouped_Class_comp_nmds_weighted$layers[-1]
ps_merged_V4_grouped_Class_comp_nmds_weighted$layers<-ps_merged_V4_grouped_Class_comp_nmds_weighted$layers[-1]
ps_merged_V4_grouped_Class_comp_nmds_weighted
ggsave("Plots/ps_merged_V4_grouped_Class_comp_nmds_weighted.svg",width=400,height=250,units=c("mm"))
```

## MERGE ALL V4 AND V9

```{r}
#Merge phyloseq objects
ps_merged<-merge_phyloseq(ps_PRJNA329332_V4,ps_PRJNA803180_V4,ps_PRJNA805287_V4,ps_PRJNA838732_V4,ps_MSc_Bonney_East_V9,ps_MSc_Joyce_V9,ps_MSc_Vanda_V9,ps_PRJNA744712_V9)

#How many ASVs and samples
ntaxa(ps_merged)
nsamples(ps_merged)

#Extract metadata and export into a csv
f<-paste(getwd(),"ps_merged_metadata.csv",sep="/")
merged_metadata<-meta(ps_merged)
#write.csv(merged_metadata,file=f)

#Compositional transformation
ps_merged_comp<-microbiome::transform(ps_merged,"compositional")
#Melt into dataframe
ps_merged_comp_melt<-psmelt(ps_merged_comp)

#Name colour palettes using this dataframe
site_palette<-unname(createPalette(length(unique(ps_merged_comp_melt$env_broad_scale)),c("#010101","#ffffff")))
```

### Averaging across replicates

```{r}
#Convert NAs to 0 abundance
otu_table(ps_merged)[is.na(otu_table(ps_merged))]<-0

#Preserve metadata
metadata_backup <- as(sample_data(ps_merged_comp), "data.frame")
metadata_backup$SampleID <- rownames(metadata_backup)

#Merge samples according to their replicate group
ps_merged_grouped<-merge_samples(ps_merged,group="RepGroup")


#Create new metadata manually, using first row per group since the replicates had identical metadata
metadata_grouped <- metadata_backup %>%
  group_by(RepGroup) %>%
  slice(1) %>%
  ungroup()

#Force setting of sample names to correct rownames (RepGroup)
metadata_grouped_sd<-sample_data(as.data.frame(metadata_grouped))
sample_names(metadata_grouped_sd)<-metadata_grouped$RepGroup
sample_data(ps_merged_grouped) <- metadata_grouped_sd


#get number of replicates per group
rep_counts<-table(sample_data(ps_merged)$RepGroup)

#extract and convert otu table to a matrix
otu_mat<-as(otu_table(ps_merged_grouped),"matrix")
otu_mat<-t(otu_mat)
#ensure sample names match
missing_names <- setdiff(colnames(otu_mat), names(rep_counts))
if (length(missing_names) > 0) {
  print("Warning: These merged sample names are missing from rep_counts:")
  print(missing_names)
}
colnames(otu_mat) <- as.character(colnames(otu_mat))
names(rep_counts) <- as.character(names(rep_counts))


#Average
otu_mat<-sweep(otu_mat,2,rep_counts[colnames(otu_mat)],"/")
otu_mat[is.na(otu_mat)] <- 0

#Check and align sample names
otu_mat <- otu_mat[, rownames(sample_data(ps_merged_grouped))]

#Return back to phyloseq object
otu_table(ps_merged_grouped) <- otu_table(otu_mat, taxa_are_rows = TRUE)
```

## Relative abundance plots

### Class level

```{r}
#Class level
ps_merged_grouped_Class_comp<-process_ps_18S(ps_merged_grouped,glom="Class",transformation="comp",remove_na=F)
#Extract taxonomy table
tax<-tax_table(ps_merged_grouped_Class_comp)
unresolved_taxa<-is.na(tax[,"Class"])
#Count number of unresolved asvs
num_unresolved<-sum(unresolved_taxa)
cat("Number of ASVs not resolved to class level:",num_unresolved,"\n")
#Make a copy to edit
tax_fixed<-tax
#Make a new pseudo-class called 'Unresolved', and apply it to higher taxonomic levels
tax_fixed[unresolved_taxa,"Class"]<-"Unresolved"
tax_fixed[unresolved_taxa,"Subdivision"]<-"Unresolved"
tax_fixed[unresolved_taxa,"Division"]<-"Unresolved"
tax_fixed[unresolved_taxa,"Supergroup"]<-"Unresolved"
tax_fixed[unresolved_taxa,"Domain"]<-"Unresolved"
#Replace tax table of phyloseq object
tax_table(ps_merged_grouped_Class_comp)<-tax_fixed
#Reprocess the ps object
ps_merged_grouped_Class_comp<-tax_glom(ps_merged_grouped_Class_comp,taxrank="Class")
ps_merged_grouped_Class_comp_melt<-psmelt(ps_merged_grouped_Class_comp)
ps_merged_grouped_Class_comp_melt[["Class"]]<-as.character(ps_merged_grouped_Class_comp_melt[["Class"]])
ps_merged_grouped_Class_comp_melt<-prune_0(ps_merged_grouped_Class_comp_melt)

#Reordering
reordering_df<-ps_merged_grouped_Class_comp_melt
reordering_df$RepGroup<-fct_relevel(reordering_df$RepGroup,"WH_littoral_mat","WH_shallow_mat","WH_deep_mat","WH_snow_pack","WH_water_track","WH_lake_outlet","WH_epilimnion","WH_mixolimnion","WH_hypolimnion","Sheldrake_Lake","Sheldrake_River_Tributary_1","Sheldrake_River_Tributary_2","Sheldrake_River_1","Sheldrake_River_2","Sheldrake_River_3","Sheldrake_River_4","Sheldrake_River_5","Sheldrake_River_6","Sheldrake_River_7","Sheldrake_River_8","Sheldrake_River_9","Hudson_Bay_Estuary_1","Hudson_Bay_Estuary_2","Hudson_Bay_Estuary_3","Hudson_Bay","WS_surface","KS_surface","KS_upper_0.5","KS_upper_1","KS_mid_1.5","KS_mid_2","KS_chemocline_2.5","KS_chemocline_3","KS_chemocline_3.1","KS_anoxic_3.5","KS_anoxic_4","Chang_Hu_17","Chang_Hu_18","Chang_Hu_19","Kitec_17","Kitec_18","Kitec_19","Xi_Hu_17","Xi_Hu_18","Xi_Hu_19","Yan_Ou_17","Yan_Ou_18","Yan_Ou_19","Yue_Ya_17","Yue_Ya_18","Yue_Ya_19","CAN47","CAN61","CAN64","CAN75","CAN81","CAN87","CAN101","CAN102","CAN111","TAY18","TAY19","TAY29","TAY38","TAY40","TAY42","TAY44","COM7","COM54","COM55","COM57","COM58","COM62","COM65","COM80","COM81","DGC4","Ust4","MVCS3","MVC4","MVC5","MVC6","MVC7","LWC1","LWC2","LWC3","LWC4","LWC5","LKGC1","LKGC2","LKGC3","LKGC4","LKGC5","KSC13","KSC14","KSC16","KSC17","KSC23","KSC33","SX","SY","SZ","SvA","SvB","SvC","SvE","IM1","IM2","IM3","IM4","IC1","IC2","GrC8","GrC11","GrE8","Bonney_East_18","Bonney_East_22","Joyce_25","Joyce_27","Joyce_37","Vanda_63","Vanda_66","Vanda_71")
reordering_df$env_broad_scale<-fct_relevel(reordering_df$env_broad_scale,"Ward_Hunt_Lake","Sheldrake_Lake","Sheldrake_River_Tributary","Sheldrake_River","Hudson_Bay_Estuary","Hudson_Bay","White Sea","Kislo-Sladkoe Lake","Chang_Hu_Lake","Kitec_Lake","Xi_Hu_Lake","Yan_Ou_Lake","Yue_Ya_Lake","Canada_glacier","Taylor_glacier","Commonwealth_glacier","Diamond_glacier","Utsteinen_nunatak","Miers_glacier","Lower_Wright_glacier","Lower_Koettlitz_glacier","Upper_Koettlitz_glacier","Storglaciaren","Midtre_Lovenbreen_glacier","Greenland_Ice_margin_glacier","Greenland_Ice_core_glacier","Kangerlussuaq_glacier","Lake_Bonney_East","Lake_Joyce","Lake_Vanda")
ps_merged_grouped_Class_comp_melt<-reordering_df


# Calculate mean abundance by Class
mean_abund <- ps_merged_grouped_Class_comp_melt %>%
  group_by(Class) %>%
  summarise(mean_abund = mean(Abundance, na.rm = TRUE)) %>%
  arrange(desc(mean_abund))
# Get levels in abundance order
Class_levels<-mean_abund$Class
# Move 'Unresolved' to the end
Class_levels<-c(setdiff(Class_levels, "Unresolved"), "Unresolved")
# Apply new levels
ps_merged_grouped_Class_comp_melt$Class<-factor(ps_merged_grouped_Class_comp_melt$Class,levels=Class_levels)

#Group low abundance data together (<0.01%)
ps_merged_grouped_Class_comp_melt[["Class"]][ps_merged_grouped_Class_comp_melt$Abundance<0.001]<-"<0.1% total abundance"
#Remove any NAs generated
ps_merged_grouped_Class_comp_melt$Class[is.na(ps_merged_grouped_Class_comp_melt$Class)] <- "Unresolved"
#Palette with lots of colours
site_palette<-unname(createPalette(length(unique(reordering_df$env_broad_scale)),c("#010101","#ffffff")))

#plot bubble plot
bubble_plot<-ggplot(ps_merged_grouped_Class_comp_melt,aes(shape=geo_loc_name,x=RepGroup,y=fct_rev(Class)))+
  geom_point(aes(size=Abundance,color=as.factor(env_broad_scale),shape=env_medium),alpha=0.8)+
  scale_size_continuous("Relative Abundance",limits=c(0.00001,1),range=c(1.5,15),breaks=c(0.10,0.25,0.50,0.75,1),labels=c("10%","25%","50%","75%","100%"))+
  scale_color_manual("Location",values=site_palette)+
  scale_shape_manual("Country",values=c(16,15,17,18))+
  labs(y="Class",x="Sample")+
  theme_bubble()+
  #expand_limits(y=c(0,length(levels(ps_merged_grouped_Class_comp_melt$"Class"))),x=c(0,length(levels(ps_merged_grouped_Class_comp_melt$"RepGroup"))))+
  guides(size=guide_legend(order=1),colour=guide_legend(order=2))
bubble_plot$layers
bubble_plot
ggsave("Plots/ps_merged_grouped_Class_bubble_plot_with_unresolved.svg",width=700,height=600,units=c("mm"))
```

## Beta diversity - NMDS

```{r}
#NMDS
ps_merged_grouped_Class_comp<-process_ps_18S(ps_merged_grouped,glom="Class",transformation="comp",remove_na=T)
ps_merged_grouped_Class_comp_melt<-psmelt(ps_merged_grouped_Class_comp)
ps_merged_grouped_Class_comp_melt<-prune_0(ps_merged_grouped_Class_comp_melt)
#Reordering
#Reordering
reordering_df<-ps_merged_grouped_Class_comp_melt
reordering_df$RepGroup<-fct_relevel(reordering_df$RepGroup,"WH_littoral_mat","WH_shallow_mat","WH_deep_mat","WH_snow_pack","WH_water_track","WH_lake_outlet","WH_epilimnion","WH_mixolimnion","WH_hypolimnion","Sheldrake_Lake","Sheldrake_River_Tributary_1","Sheldrake_River_Tributary_2","Sheldrake_River_1","Sheldrake_River_2","Sheldrake_River_3","Sheldrake_River_4","Sheldrake_River_5","Sheldrake_River_6","Sheldrake_River_7","Sheldrake_River_8","Sheldrake_River_9","Hudson_Bay_Estuary_1","Hudson_Bay_Estuary_2","Hudson_Bay_Estuary_3","Hudson_Bay","WS_surface","KS_surface","KS_upper_0.5","KS_upper_1","KS_mid_1.5","KS_mid_2","KS_chemocline_2.5","KS_chemocline_3","KS_chemocline_3.1","KS_anoxic_3.5","KS_anoxic_4","Chang_Hu_17","Chang_Hu_18","Chang_Hu_19","Kitec_17","Kitec_18","Kitec_19","Xi_Hu_17","Xi_Hu_18","Xi_Hu_19","Yan_Ou_17","Yan_Ou_18","Yan_Ou_19","Yue_Ya_17","Yue_Ya_18","Yue_Ya_19","CAN47","CAN61","CAN64","CAN75","CAN81","CAN87","CAN101","CAN102","CAN111","TAY18","TAY19","TAY29","TAY38","TAY40","TAY42","TAY44","COM7","COM54","COM55","COM57","COM58","COM62","COM65","COM80","COM81","DGC4","Ust4","MVCS3","MVC4","MVC5","MVC6","MVC7","LWC1","LWC2","LWC3","LWC4","LWC5","LKGC1","LKGC2","LKGC3","LKGC4","LKGC5","KSC13","KSC14","KSC16","KSC17","KSC23","KSC33","SX","SY","SZ","SvA","SvB","SvC","SvE","IM1","IM2","IM3","IM4","IC1","IC2","GrC8","GrC11","GrE8","Bonney_East_18","Bonney_East_22","Joyce_25","Joyce_27","Joyce_37","Vanda_63","Vanda_66","Vanda_71")
reordering_df$env_broad_scale<-fct_relevel(reordering_df$env_broad_scale,"Ward_Hunt_Lake","Sheldrake_Lake","Sheldrake_River_Tributary","Sheldrake_River","Hudson_Bay_Estuary","Hudson_Bay","White Sea","Kislo-Sladkoe Lake","Chang_Hu_Lake","Kitec_Lake","Xi_Hu_Lake","Yan_Ou_Lake","Yue_Ya_Lake","Canada_glacier","Taylor_glacier","Commonwealth_glacier","Diamond_glacier","Utsteinen_nunatak","Miers_glacier","Lower_Wright_glacier","Lower_Koettlitz_glacier","Upper_Koettlitz_glacier","Storglaciaren","Midtre_Lovenbreen_glacier","Greenland_Ice_margin_glacier","Greenland_Ice_core_glacier","Kangerlussuaq_glacier","Lake_Bonney_East","Lake_Joyce","Lake_Vanda")
ps_merged_grouped_Class_comp_melt<-reordering_df

#colour palette
site_palette<-unname(createPalette(length(unique(reordering_df$env_broad_scale)),c("#010101","#ffffff")))

#Ordination
ps_merged_grouped_Class_comp_ord_weighted<-ordinate(ps_merged_grouped_Class_comp,"NMDS","bray")
#Stress: 0.1589103
#ANOSIMs
ps_merged_grouped_Class_comp_loc_group=get_variable(ps_merged_grouped_Class_comp, "env_broad_scale")
ps_merged_grouped_Class_comp_loc_anosim=anosim(distance(ps_merged_grouped_Class_comp, "bray"), ps_merged_grouped_Class_comp_loc_group)
ps_merged_grouped_Class_comp_loc_anosim
#ANOSIM: Stress = 0.1447421, ANOSIM R-Stat 0.4238, p value = 0.001


ps_merged_grouped_Class_comp_dataframe<-as.data.frame(sample_data(ps_merged_grouped_Class_comp))
ps_merged_grouped_Class_comp_nmds_weighted<-plot_ordination(ps_merged_grouped_Class_comp,ps_merged_grouped_Class_comp_ord_weighted,type="samples",
                                      color="env_broad_scale",label="RepGroup",shape="env_medium")+
  ggtitle("ANOSIM: Stress = 0.1447421, ANOSIM R-Stat 0.4238, p value = 0.001")+
  scale_colour_manual("Location",values=site_palette)+
  scale_shape_manual("Media",values=c(15,16,17))+
  geom_point(size=4,alpha=0.8)+
  geom_text_repel(mapping=aes(x=NMDS1,y=NMDS2,label=RepGroup,fontface="bold"),inherit.aes=F,,size = 3,vjust=0,nudge_y=0.04,max.overlaps=100)+
  #geom_mark_hull(aes(fill=env_broad_scale,colour=env_broad_scale),show.legend=F)+
  expand_limits(x=1.5,y=1)+
  theme_mine()
ps_merged_grouped_Class_comp_nmds_weighted$layers
ps_merged_grouped_Class_comp_nmds_weighted$layers<-ps_merged_grouped_Class_comp_nmds_weighted$layers[-1]
ps_merged_grouped_Class_comp_nmds_weighted$layers<-ps_merged_grouped_Class_comp_nmds_weighted$layers[-1]
ps_merged_grouped_Class_comp_nmds_weighted
ggsave("Plots/ps_merged_grouped_Class_comp_nmds_weighted.svg",width=400,height=250,units=c("mm"))
```

## Import datasets which used the same primer, processed identically and merged in qiime

V9_1391f_EukBr

Trimming -\> trim left: 5; truncate forward: 200; truncate reverse: 185

```{r}
#Read in OTU table
V9_1391f_EukBr_otu_table=read.csv("data/V9_1391f_EukBr/V9_1391f_EukBr_table.csv",sep=",",row.names=1,na.strings="")
V9_1391f_EukBr_otu_table=as.matrix(V9_1391f_EukBr_otu_table)

#Read in taxonomy
V9_1391f_EukBr_taxonomy=read.csv("data/V9_1391f_EukBr/V9_1391f_EukBr-pr2-taxonomy.csv",sep=",",row.names=1,na.strings="")
V9_1391f_EukBr_taxonomy_df<-as.data.frame(V9_1391f_EukBr_taxonomy)
#Filter out low confidence assignments
V9_1391f_EukBr_taxonomy_97_df<-V9_1391f_EukBr_taxonomy_df%>%filter(as.numeric(Confidence)>=0.97)
V9_1391f_EukBr_taxonomy_97=as.matrix(V9_1391f_EukBr_taxonomy_97_df)

#Read in metadata
V9_1391f_EukBr_metadata=read.csv("data/V9_1391f_EukBr/V9_1391f_EukBr_metadata.csv",row.names=1,na.strings="")

#Read in tree
phy_tree_V9_1391f_EukBr<-read.tree("data/V9_1391f_EukBr/V9_1391f_EukBr_unrooted_tree.nwk")

#Import as PhyloSeq objects
OTU_V9_1391f_EukBr=otu_table(V9_1391f_EukBr_otu_table,taxa_are_rows=TRUE)
TAX_V9_1391f_EukBr=tax_table(V9_1391f_EukBr_taxonomy_97)
META_V9_1391f_EukBr=sample_data(V9_1391f_EukBr_metadata)

#Check sample names are identical and same order
sample_names(OTU_V9_1391f_EukBr)
sample_names(META_V9_1391f_EukBr)

#Merge into one PhyloSeq object
ps_V9_1391f_EukBr=phyloseq(OTU_V9_1391f_EukBr,TAX_V9_1391f_EukBr,META_V9_1391f_EukBr,phy_tree_V9_1391f_EukBr)
ps_V9_1391f_EukBr

#Filter out any ASVs not assigned further than 'Eukaryota' by removing any with 'NA' in the next taxonomic rank (Supergroup)
keep_taxa<-taxa_names(ps_V9_1391f_EukBr)[!is.na(tax_table(ps_V9_1391f_EukBr)[,"Supergroup"])]
ps_V9_1391f_EukBr<-prune_taxa(keep_taxa,ps_V9_1391f_EukBr)

#Add prefix to front of ASVs in ASV table and taxonomy table
prefix<-"V9_1391f_EukBr_"
old_names<-taxa_names(ps_V9_1391f_EukBr)
new_names<-paste0(prefix,old_names)
taxa_names(ps_V9_1391f_EukBr)<-new_names
otu_table(ps_V9_1391f_EukBr)<-otu_table(ps_V9_1391f_EukBr)[new_names,]
tax_table(ps_V9_1391f_EukBr)<-tax_table(ps_V9_1391f_EukBr)[new_names,]
```
